---
title: "FE515_2022A_Lecture_2"
author: "Cheng Lu"
output: pdf_document
---

Learning objectives:
  1. Data Frame
  2. Import CSV file
  3. Missing Values
  4. User defined functions
  5. Vectorized Operation*
  6. Apply family


# 1. Data Frame
```{r}
# two vectors
gender <- c("boy", "girl", 'boy', 'girl')
ages <- c(11, 12, 10, 13)

# check data type
typeof(gender)
typeof(ages)

# create a dataframe
df <- data.frame(gender, ages)
df

# a special list
typeof(df)
class(df)

# data frames can store different type of objects in each colum
typeof(df$ages)
typeof(df$gender)

# strings as factors
df <- data.frame(gender, ages, stringsAsFactors = T)
df
df$gender # factors
```

## 1.1 factor*

Factor is used to categorize the data and store it as levels.
They can store both strings and integers.
they are useful in the columns which have a limited number of unique values. 
Like "Male, "Female" and True, False

```{r factor}
# encode vector as factor
data <- c("East","West","East","North","South","North","East","West","West","West","East","North")
print(data)
print(is.factor(data))

# Apply the factor function.
factor_data <- factor(data)
print(is.factor(factor_data))
print(factor_data)

# Change element values
factor_data <- factor(data, labels = c(1, 2, 3, 4))
print(factor_data)
```


# 2. Import CSV

```{r}
# read csv file
read.csv("google.csv")
GOOG <- read.csv(file = "google.csv", header = T, row.names = 1)

# current working directory
getwd()

# set working directory
path <- "/Users/cheng/Google Drive/PhD/Teaching/FE 515 R/2022 fall/code"
setwd(path)

# first several rows of GOOG
head(GOOG) 

# column names
names(GOOG)

# get open price
GOOG$GOOG.Open

# get adjusted close price
GOOG$Adj.Close

# GOOG is a data frame
class(GOOG) 
typeof(GOOG)
```

# 3. Missing Values
```{r}
# NA
# Not available
head(GOOG)

# NaN
# Not a number
x <- c(-1,0,1)
x/0


# identify NA and NaN
# nan is a special case of NA
is.na(c(1,NA,NaN))
is.nan(c(1,NA,NaN))

# working with NA
x <- c(1, 2, 3, NA, NA, 6, NA, 8)
x
x.na <- is.na(x) 

# method 1
x[x.na]
!x.na
x[!x.na]

# method 2
?na.omit
y <- na.omit(x)
as.numeric(y)
```


# 4. User defined functions

see slides for syntax

function without argument

```{r}
myfunc1 <- function(){
  a <- c(1,2,3)
  a.square.root <- sqrt(a)
  return(a.square.root)
}

result <- myfunc1()
print(result)
myfunc1()# error: unused argument
```

function with arguments

```{r}
myfunc2 <- function(arg1, arg2){
  a <- arg1 * arg2
  return(a)
}

myfunc2(2)

result <- myfunc2(c,c) # call the new function
print(result)
myfunc2(c)# Error in myfunc2() : argument "arg1" is missing, with no default
```

function with default arguments

```{r}
get.circle.area <- function(radius = 1, pi = 3.14){ 
  a <- pi * radius ^ 2
  return(a)
}

get.circle.area(radius = 6)
get.circle.area(3, 3)
get.circle.area()
get.circle.area(5)# first argument is replaced by 5
```

Functions can use global variables

global variables
local variables

```{r}
var <- "Hello World!"

print.sth <- function(){
 print(var)
  var <<- "Hello R World!" # how about "<<-" rather than "<-" ?
 print(var)
}

print.sth()
var
```

# 5. Vectorized Operation*
```{r}
x <- 1:4
y <- 6:9

# want to do x + y
# method 1: using for loop
result <- c()
for (i in 1:4) {
  result[i] <- x[i] + y[i] 
  }
result

# method 2: vectorized operation
result <- x + y
result
```

Another example

```{r}
x <- 1:4
x > 2
```

Apply vectorized operation to functions defined by user

```{r}
# user defined functions 
mySquare <- function(x) { 
  return (x^2)
}

# vector as example
x <- 1:4
y <- 6:9

# vectorized operation
mySquare(x)
mySquare(y)
```

# 6. Apply Family


## 6.1 apply

apply(x,margin,function): Returns a vector or array or list of values 
obtained by applying a function to margins of an array or matrix. 
Margin is the name of the dataset used.
1 indicates rows, 2 indicates columns, 3 indicates matrix

```{r}
# create an example
M <- matrix(c(1:9), byrow = T, nrow = 3)
print(M)

# apply
result <- apply(X = M, MARGIN = 1, FUN = sum) # get sum of elements for matrix 
print(result)
```

## 6.2 lapply

'lapply' works with list (takes list as input and return a list)

```{r}
# take matrix as input return list
lapply(M, sum)

# takes list as input and return a list
x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
print(x)
lapply(x, mean)
```

## 6.3 sapply

'sapply' is a simple version (wrapper) of 'lapply' (work with list and return vector)

```{r}
# takes list as input and return a vector
sapply(x, mean)

# take matrix as input return vector
col.mean <- sapply(M, mean)
class(col.mean)
```



