---
title: "FE515_Lecture_1"
author: "Cheng Lu"
date: "9/11/2022"
output: pdf_document
---

Learning objectives:
  1. Atomic Class
  2. Explicit Coercion
  3. Vector
  4. Matrix
  5. List 
  6. Subsetting 
  7. Operator
  8. Loop
  
# 1. Atomic Class

First thing

```{r}
# say hello to R!

```


```{r}
# 1. numeric 
a <- 1
b <- 2
c <-  a + b
print(c)
mode(c)

# 2. integer
a <- 41L
print(a)
typeof(a)

# 3. logical
a <- TRUE
b <- FALSE
print(a)
mode(a)

# 4. character
a <- "Hello world!"
print(a)
mode(a)

# 5. complex number
a <- 3 + 4i
b <- 2 + 2i
c <- a - b
print(c)
mode(c)
```

# 2. Explicit Coercion (transformation)

```{r}
x <- 

# Convert an object to numeric
as.numeric(x)

# Convert an object to logical
as.logical(x)

# Convert an object to character
as.character(x)
```

# 3. Vector

$$
x = \left(\begin{array}{c} 
1 \\
2 \\
3 
\end{array}\right)
$$ 

Creating Vectors

```{r}
x <- c(4, 5, 6) # numeric

x <- c(TRUE, FALSE) # logical
x <- c(T, F) # logical
x <- c('a', 'b', 'c') # character
x <- 1:100 # integer
x <- c(1+0i, 3+5i) # complex
```

Other ways to create vectors

```{r}
x <- seq(10, 20, by = 2)
x

y <- rep(x = c(1, 2, 3), 2) # this "x" is different from the former x
y

z <- c(x, 0, y)
z

a <- 1:length(x)+1 # ’:’ has higher precedence than ’+’
a
```

vector calculation

```{r}
x <- 1:5

# plus
2*x+1

# minus
x - 5

# multiply
x * 3

# divide
x / 3

# power
x^2

# modulo
5 %% 3
x %% 3

# inner product
y <- rep(1, 5)
x%*%y
1*1 + 2*1 + 3*1 + 4*1 + 5*1
```

# 4. Matrix

$$
A = \left(\begin{array}{ccc} 
1 & 2 & 3\\
4 & 5 & 6\\
7 & 8 & 9
\end{array}\right)
$$ 

Creating matrix

```{r}
x <- 1:9
A <- matrix(data=x, nrow=3, ncol=3, byrow=T)

nrow(A)
ncol(A)
t(A)
```

Other ways to create matrix

```{r}
x <- 1:3
y <- 4:6
rbind(x, y)
cbind(x, y)
```

Matrix Operations*

```{r}
# transpose of A
t(A)

# determinant of A
det(A) 

# eigenvalues of A
eigen(A)$values

# eigenvectors of A
eigen(A)$vectors 

# element-wise multiplication
A * B 

# matrix multiplication
A %*% B 

# inverse of matrix A
solve(A) 

# solve linear equation Ax = b
solve(A, b) 
```

Other Operations for Vector or Matrix*

```{r}
# summation of all elements in A
sum(A)

# product of all elements in A
prod(A)

# maximum element in A
max(A)

# minimum element in A
min(A)

# exponential of each element in A
exp(A)

# logarithm of each element in A
log(A)

# absolute value of each element in A
abs(A)
```

# 5. List

```{r}
l <- list("John", 12345, "Male")
l

l <- list(name = "John", ID = 12345, gender = "Male")
l$name
```

# 6. Subsetting (indexing)

```{r}
# vector
x <- 1:10
x[3]
x[c(3,7)]
x[-2]
x[-c(2:5)]

# matrix
m <- matrix(x, nrow = 2, byrow = T)
m
m[2,2]
m[2, c(1,3)]
m[,c(3,4)]
m[,-c(2:4)]

# list
y <- c('a', 'b', 'c')
l <- list(numbers = x, chars = y)
l$numbers[3]


l[[1]][3]
l[[1]]
l$chars[c(1,3,4)]
```

Assign Values

```{r}
a <- x[c(2:5)]
a

m[,4:5] <- 0
m

l$numbers[-c(2,3,4)] <- 0
l$numbers
l$chars
```


# 7. Operator

## 7.1 Arithmetic Operations

example 1:

$$
d_{1}=\frac{\ln \frac{S_{0}}{K}+\left(r+\frac{1}{2} \sigma^{2}\right) T_{1}}{\sigma \sqrt{T_{1}}}
$$
```{r}
S0 <- 100
K <- 100
T1 <- 1
sigma <- 0.2
r <- 0.05

d1 <- (log(S0/K)+(r+0.5*sigma^2)*T1)/(sigma*sqrt(T1))
d1
```

example 2:

$$
\frac{1}{\sqrt{2 \pi}} e^{-\frac{x^{2}}{2}}
$$
```{r}
x <- 0
1/sqrt(2*pi)*exp(-x^2/2)

x <- -3:3
1/sqrt(2*pi)*exp(-x^2/2)
```

## 7.2 Logical Operations

```{r}
a <- pi # a = 3.1415926...
a > 3

# equal
y <- 1:7
y
y == 3

# not equal
y != 3

# and
a > 3 & a < 4

# or
a > 3 | a < 4
```

# 8. if else statement

```{r}
x <- T

if (x){
  "x is true"
} else {
  "x is false"
}
```

another example

```{r}
x <- c(1, 2, 3)

print(x)

if(2 %in% x) {
   print("2 is found")
} else {
   print("2 is not found")
}
```


# 9. Loop

repeat loop

```{r}
a <- c("Hello")
num <- 1

repeat {
  print(a)
  num = num + 1
  if(num > 5){ break } # When the break is inside a loop, the loop is immediately terminated, and...
  # goes to the next statement following the loop
}
```

while loop

```{r}
i <- 1
while(i <= 100){ # infinite loop, press 'stop' to stop printing
  print(i)
  i <- i + 1
}
```

for loop

```{r}
for (i in 1:5) {
  print(i)
}


sums <- 0
for (i in 1:100) {
  sums <- sums + i
}
sums
```

nested for loop*

$$
A = \left(\begin{array}{ccc} 
1 & 2 & 3\\
4 & 5 & 6\\
7 & 8 & 9
\end{array}\right)
$$ 


```{r}
A2 <- matrix(NA, nrow = 3, ncol = 4) # NA matrix
A1 <- matrix(1:12, nrow = 3, byrow = T)

for (i in 1:3) {# i is row index
  for (j in 1:4) {# j is col index
    A2[i, j] <- A1[i, j] * 10
  }
}

A2
```








