---
title: "FE515_2022A_Lecture_9"
author: "Cheng Lu"
output: pdf_document
---
Learning objectives:
  1. Black-Scholes Model
  2. Implied Volatility
    2.1 Implied Volatility for Market Data

# 1. Black-Scholes Model

(see slides for details)

the call option price c at time 0 

$$
\begin{align}
c\left(S_{0}, K, T, \sigma, r\right)&=e^{-r T} E^{Q}\left[(S(T)-K)_{+}\right]\\
 &=S_{0} N\left(d_{1}\right)-e^{-r T} K N\left(d_{2}\right)\\
where\\
d_{1}&=\frac{\ln \frac{S_{0}}{K}+\left(r+\frac{1}{2} \sigma^{2}\right) T}{\sigma \sqrt{T}}\\
d_{2}&=\frac{\ln \frac{S_{0}}{K}+\left(r-\frac{1}{2} \sigma^{2}\right) T}{\sigma \sqrt{T}}\\
 &=d_{1}-\sigma \sqrt{T}
\end{align}
$$

```{r}
# Hyper-parameter
S0 <- 100 # spot price
K <- 100 # strike price
T1 <- 1 # maturity # if we let T <- 1, "T" would not be "TRUE" anymore
sigma <- 0.2 # volatility
r <- 0.05 # risk free rate

# BS formula
d1 <- (log(S0/K) + (r+0.5*sigma^2)*T1)/(sigma*sqrt(T1))
d2 <- d1 - sigma*sqrt(T1)
(c <- S0*pnorm(d1) - exp(-r*T1)*K*pnorm(d2))

```

Wrap into a function

```{r}
bs.call <- function(S0, K, T1, sigma, r){
d1 <- (log(S0/K) + (r+0.5*sigma^2)*T1)/(sigma*sqrt(T1))
d2 <- d1 - sigma*sqrt(T1)
S0*pnorm(d1) - exp(-r*T1)*K*pnorm(d2)
# return(S0*pnorm(d1) - exp(-r*T1)*K*pnorm(d2))
}
```

Test the function

```{r}
# case 1
bs.call(S0, K, T1, sigma, r) # bs.call(S0, K, T1, r, sigma) doesn’t work

# case 2
bs.call(100, 100, 1, r = 0.05, sigma = 0.2)

# case 3
bs.call(100, 100, 1, 0.2, 0.05)
```

# 2. Implied Volatility

We can use the root finding methods (e.g. bisection method in L8) to find $\sigma$

$$
c\left(S_{0}, K, T, \sigma, r\right)=P \Leftrightarrow c\left(S_{0}, K, T, \sigma, r\right)-P=0
$$

bisection method to find sigma

```{r}
# form the equation
price.diff <- function(sigma) bs.call(S0, K, T1, sigma, r) - 10

# bisection method
bisection.new(price.diff, 0, 1)
```

newton's method to find sigma

$$
\begin{align}
\text { Vega }&=\frac{\partial c}{\partial \sigma}=\sqrt{T} S_{0} N^{\prime}\left(d_{1}\right)\\
d_{1}&=\frac{\ln \frac{S_{0}}{K}+\left(r+\frac{1}{2} \sigma^{2}\right) T}{\sigma \sqrt{T}}
\end{align}
$$

```{r}
# derived function
Vega <- function(S0, K, T1, sigma, r){
d1 <- (log(S0/K) + (r+0.5*sigma^2)*T1)/(sigma*sqrt(T1))
sqrt(T1)*S0*dnorm(d1)
}
dprice.diff <- function(sigma) Vega(S0,K,T1,sigma,r)

# newton's method
Newton_Raphson(price.diff,dprice.diff,0.25)
```

Wrap up everything into a single function

```{r}
# integration
implied.vol.call <- function(S0, K, T1, r, price, method="bisection"){
  
  price.diff <- function(sigma)bs.call(S0, K, T1, sigma, r) - price
  
  if(method == "bisection"){
    return(bisection.new(price.diff, 0.01, 5))
  }else if(method == "Newton-Raphson"){
    dprice.diff <- function(sigma)Vega(S0, K, T1, sigma, r)
    return(Newton_Raphson(price.diff, dprice.diff, 0.25))
  }
}

# test cases
implied.vol.call(S0,K,T1,r,10)

implied.vol.call(S0,K,T1,r,10,"Newton-Raphson")

```

## 2.1 Implied Volatility for Market Data

get market data

```{r}
library(quantmod)

# underlying asset price S_0
(SPY.S0 <- getQuote("SPY")$Last)

# risk free rate
# FRED | FAMA FRENCH DATA
r <- 0.07 * 0.01
```

get option data

```{r}
SPY.option <- getOptionChain("SPY")
head(SPY.option$calls)# nearest maturity option
```

determine parameters value

```{r}
# strike price
(SPY.K1 <- SPY.option$calls$Strike[1])

# option price
(SPY.Price1 <- 0.5*(SPY.option$calls$Bid[1] + SPY.option$calls$Ask[1]))

# time to maturity
(T1 <- as.numeric(as.Date("2022-07-27") - Sys.Date())/252)
```

calculate implied volatility

```{r}
implied.vol.call(SPY.S0,SPY.K1,T1,r,SPY.Price1)
```

modify bisection method for all option contract

```{r}
bisection.new <- function(f, a, b, tol = 0.001, N.max = 100){
  # assign value
  f.a <- f(a)
  f.b <- f(b)
  
  # initial check
  if(is.na(f.a*f.b) || f.a*f.b > 0){# only modified this part
    return(NA)
  }else if(f.a == 0){
    return(a)
  }else if(f.b == 0){
    return(b)
  }
  
  # the same searching process
  for(n in 1:N.max){
    c <- (a+b)/2
    f.c <- f(c) # call function 'f' once each iteration
    if(f.c == 0 || abs(b - a) < tol){ # check absolute value
      break
    }
    if(f.a*f.c < 0){
      b <- c
      f.b <- f.c
    }else{
      a <- c
      f.a <- f.c
    }
  }
  
  return(c) # return a value rather than print it out
}
```

implied volatility for all contracts (volatility smile)

```{r}
SPY.call.strike <- SPY.option$calls$Strike
SPY.call.price <- 0.5*(SPY.option$calls$Bid+SPY.option$calls$Ask)
sigma.call <- NULL
for (i in 1:nrow(SPY.option$calls)){
sigma.call[i] <- implied.vol.call(SPY.S0,SPY.call.strike[i],
T1, r, SPY.call.price[i])
}
plot(SPY.call.strike, sigma.call, type = "l")
```


