---
title: "FE515_2022A_Lecture_10"
author: "Cheng Lu"
output: pdf_document
---
Learning objectives:
  1. Monte-Carlo Simulation
  2. Improve Efficiency for MC Simulation

# 1. Monte-Carlo Simulation

Euler discretization for stock price

$$
\begin{aligned}
S_{i+1} &=S_{i}+r S_{i} h+\sigma S_{i}\left(W_{i+1}^{Q}-W_{i}^{Q}\right) \\
&=S_{i}+r S_{i} h+\sigma S_{i} \sqrt{h} Z_{i},\;\; i=0, \ldots, n-1\\
\text{where}&\\
&\sqrt{h} Z_{i}=W_{i+1}^{Q}-W_{i}^{Q} \sim N(0, h)\\
&Z_{i} \sim N(0,1)
\end{aligned}
$$

implementation of Euler discretization

```{r}
set.seed(2)
# hyper-parameter
S0 <- 100
K <- 100
T1 <- 1
sigma <- 0.2
r <- 0.05
n <- 252 # number of steps
h <- T1/n # one day time difference
h
S <- c() # same as "S <- NULL" 
Z <- c()
S[1] <- S0 # start at S[1], since S[0] not available in R

# stock price at each time step
for (i in 1:n) {
  Z[i] <- rnorm(1)
  S[i+1] <- S[i] + r*S[i]*h + sigma*S[i]*Z[i]*sqrt(h)
}

# visualization
t <- seq(from = 0, to = T1, by = h)
plot(t, S, type = "l")
```

Theoritical expression for stock price

$$
\begin{gathered}
S\left(t_{i}\right)=S_{0} \exp \left(\left(r-\frac{1}{2} \sigma^{2}\right) t_{i}+\sigma W\left(t_{i}\right)\right) \\
W\left(t_{i}\right)=W_{i}=\sum_{j=0}^{i-1}\left(W_{j+1}-W_{j}\right)=\sum_{j=0}^{i-1} \sqrt{h} Z_{j}
\end{gathered}
$$

implementation of theoritical expression

```{r}
W <- c(0, cumsum(sqrt(h)*Z))
S.explicit <- S0*exp((r - 0.5*sigma^2)*t + sigma*W)
plot(t, S.explicit, col = "red", type = 'l')
```

compare theoritical expression with Euler discretization

```{r}
# Euler discretization
S[n+1]

# theoritical expression
S.explicit[n+1]
```

Monte-Carlo Simulation for option price

```{r}
m <- 10000
S.mat <- NULL

# simulate for m possible stock price evolutionary path
for(j in 1:m){
  for (i in 1:n) {
    Z[i] <- rnorm(1)
    S[i+1] <- S[i] + r*S[i]*h + sigma*S[i]*Z[i]*sqrt(h)
  }
  S.mat <- cbind(S.mat,S)
}

# visualization
plot(t,S.mat[,1], type = "l", main = "The First Path")
plot(t,S.mat[,2], type = "l", main = "The Second Path")
matplot(t, S.mat[,1:100], type = "l", main = "The First 100 Paths")

# calculate option price
ST <- S.mat[n+1,] # S(T) for all paths as a vector
exp(-r*T1)*mean(pmax(ST - 100, 0)) # call option price from MC simulation
bs.call(S0, 100, T1, sigma, r)# call option price from BS formula # from lecture 9
```

# 2. Improve Efficiency for MC Simulation

CPU time of running MC

```{r}
MC <- function(){
  m <- 10000
  S.mat <- NULL
  
  for (j in 1:m) {
    S[1] <- S0
    for(i in 1:n){
      Z[i] <- rnorm(1)
      S[i+1] <- S[i] + r*S[i]*h + sigma*S[i]*Z[i]*sqrt(h)
    }
    
    S.mat <- cbind(S.mat, S)
  }

  print(exp(-r*T1)*mean(pmax(S.mat[n+1,] - 100, 0)))
}

system.time(MC())
```

replace `for` loop with `replicate()` function

```{r}
MC1 <- function(){# use replicate
  m <- 10000
  one.path <- function(){
    S[1] <- S0
    
    for(i in 1:n){
      Z[i] <- rnorm(1)
      S[i+1] <- S[i] + r*S[i]*h + sigma*S[i]*Z[i]*sqrt(h)
    }
    return(S)
  }

  S.mat <- replicate(m, one.path())
  print(exp(-r*T1)*mean(pmax(S.mat[n+1,] - 100, 0)))
}

system.time(MC1())
```

More efficient when we generate random number together

```{r}
MC2 <- function(){
  m <- 10000
  one.path <- function(){
    S[1] <- S0
    Z <- rnorm(n) # take rnorm out
    for(i in 1:n){
      S[i+1] <- S[i] + r*S[i]*h + sigma*S[i]*Z[i]*sqrt(h)
    }
    return(S)
  }

  S.mat <- replicate(m, one.path())
  print(exp(-r*T1)*mean(pmax(S.mat[n+1,] - 100, 0)))
}

system.time(MC2())
```

More efficient if we update the target matrix directly

```{r}
MC3 <- function(){# initialize S and Z as matrices and update matrix
  m <- 10000
  S.mat <- matrix(0, nrow = n+1, ncol = m)
  Z <- matrix(rnorm(n*m), nrow = n)
  S.mat[1,] <- S0
  
  for (i in 1:n) {
    S.mat[i+1,] <- S.mat[i,] + r*S.mat[i,]*h + sigma*S.mat[i,]*Z[i,]*sqrt(h)
  }
  
  print(exp(-r*T1)*mean(pmax(S.mat[n+1,] - 100, 0)))
}

system.time(MC3())
```

The option payoff (S(T ) − K )+ only depends on the final price S(T)

```{r}
MC4 <- function(m){
  S.vec <- rep(S0, m)
  Z <- matrix(rnorm(n*m), nrow = n)
  for (i in 1:n) {
    S.vec <- S.vec + r*S.vec*h + sigma*S.vec*Z[i,]*sqrt(h)
  }
  print(exp(-r*T1)*mean(pmax(S.vec - 100, 0)))
}
system.time(MC4(10000))
```

