---
title: "FE515_2022A_Lecture_11"
author: "Cheng Lu"
output: pdf_document
---

Learning objectives:
  1. General Optimization
    1.1 one dimensional boxed constraint: optimize()
    1.2 multi dimensional without/with constraint: optim()
    1.3 multi dimensional linear inequality constraint: constrOptim()
  2. Linear Programming
  3. Quadratic Programming
  4. Change Problems into Standard Form

# 1. General Optimization

## 1.1 one dimensional boxed constraint: optimize()

objective function: example

$$
\min _{x \in[0,1]} f(x)=(x-a)^{2}
$$

```{r}
f <- function(x) (x - 1/3)^2 # objective function when a = 1/3
optimize(f, c(0,1))

f <- function(x, a) (x - a)^2 # objective function with parameters
optimize(f, c(0, 1), tol = 0.0000000000001, a = 1/3) # set parameters

# validate
curve(f, -2*pi, 2*pi, xname = "x")
```

## 1.2 multi dimensional without/with constraint: optim()

Rosenbrock Banana function

$$
\begin{equation}
\begin{gathered}
f\left(x_{1}, x_{2}\right)=100 *\left(x_{2}-x_{1}^{2}\right)^{2}+\left(1-x_{1}\right)^{2} \\
\left.g\left(x_{1}, x_{2}\right)=\nabla f\left(x_{1}, x_{2}\right)=\left[-400 * x_{1} *\left(x_{2}-x_{1}^{2}\right)\right)-2 *\left(1-x_{1}\right), 200 *\left(x_{2}-x_{1}^{2}\right)\right]
\end{gathered}
\end{equation}
$$


```{r}
# Rosenbrock Banana function
f <- function(x) 100*(x[2]-x[1]^2)^2 + (1-x[1])^2
optim(c(-1.2,1), f) # without gradient
```

If we use gradient g, we need to specify the method as “BFGS”, “CG” or “L-BFGS-B"

```{r}
# gradient
g <- function(x) c(-400*x[1]*(x[2]-x[1]^2)-2*(1-x[1]), 200*(x[2]-x[1]^2))
optim(c(-1.2,1), f, g, method = "L-BFGS-B")
```

If we add box constraint, we need to specify the lower bound and upper bound, and specify the method being L-BFGS-B.

```{r}
optim(c(-1.2,1), f, lower=c(-2,0), upper=c(2,2), method="L-BFGS-B")
?optim 
```

## 1.3 multi dimensional linear inequality constraint: constrOptim()

$$
\begin{equation}
\begin{gathered}
\min _{x} f(x)=0.5 x_{1}^{2}+0.5 x_{2}^{2}+0.5 x_{3}^{2}-5 x_{2} \\
\text { subject to: }-4 x_{1}-3 x_{2} \geq-8 \\
2 x_{1}+x_{2} \geq 2 \\
-2 x_{2}+x_{3} \geq 0
\end{gathered}
\end{equation}
$$

```{r}
# objective function
f <- function(x) 0.5*x[1]^2 + 0.5*x[2]^2 + 0.5*x[3]^2 - 5*x[2]
# f <- function(x)0.5*x%*%diag(3)%*%x - c(0,5,0)%*%x # matrix form

# constraint
U <- matrix(c(-4,-3,0,2,1,0,0,-2,1), 3, 3, byrow =TRUE)
c <- c(-8,2,0)

# solve the problem
constrOptim(c(2,-1,-1), f , grad = NULL, ui = U, ci = c)
```

# 2. Linear Programming

$$
\begin{equation}
\begin{gathered}
\max _{x \in \mathbb{R}^{3}} f(x)=x_{1}+9 x_{2}+x_{3} \\
\text { subject to: } x_{1}+2 x_{2}+3 x_{3} \leq 9 \\
3 x_{1}+2 x_{2}+2 x_{3} \leq 15
\end{gathered}
\end{equation}
$$


```{r}
#install.packages("lpSolve")
library(lpSolve)

# objective function
f.obj <- c(1, 9, 1) 

# constraint
f.con <- matrix (c(1, 2, 3, 3, 2, 2), nrow=2, byrow=TRUE) # constraint matrix A
f.rhs <- c(9, 15) # right hand side of the constraint b
f.dir <- c("<=", "<=") # constraint directions

# slove the problem
x <- lp ("max", f.obj, f.con, f.dir, f.rhs)
x

lp ("max", f.obj, f.con, f.dir, f.rhs)$objval # optimal objective value

lp ("max", f.obj, f.con, f.dir, f.rhs)$solution # optimal solution

# restrict x2 and x3 be integer
lp ("max", f.obj, f.con, f.dir, f.rhs, int.vec = c(2, 3))$solution
```

# 3. Quadratic Programming

$$
\begin{equation}
\begin{gathered}
\min _{x} f(x)=\frac{1}{2} x^{T} D x-d^{T} x\\
\text{subject to}: A^{T} x \geq b
\end{gathered}
\end{equation}
$$

example problem

$$
\begin{equation}
\begin{gathered}
\min _{x} f(x)=0.5 x_{1}^{2}+0.5 x_{2}^{2}+0.5 x_{3}^{2}-5 x_{2} \\
\text { subject to: }-4 x_{1}-3 x_{2} \geq-8 \\
2 x_{1}+x_{2} \geq 2 \\
-2 x_{2}+x_{3} \geq 0
\end{gathered}
\end{equation}
$$

problem statement

$$
\begin{equation}
D=\left[\begin{array}{lll}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1
\end{array}\right] ; d=\left[\begin{array}{l}
0 \\
5 \\
0
\end{array}\right] ; A^{T}=\left[\begin{array}{ccc}
-4 & -3 & 0 \\
2 & 1 & 0 \\
0 & -2 & 1
\end{array}\right] ; A=\left[\begin{array}{ccc}
-4 & 2 & 0 \\
-3 & 1 & -2 \\
0 & 0 & 1
\end{array}\right] ; b=\left[\begin{array}{c}
-8 \\
2 \\
0
\end{array}\right]
\end{equation}
$$

```{r}
#install.packages("quadprog")
library(quadprog)

# problem statement
D <- matrix(c(1,0,0,0,1,0,0,0,1),3,3) # or D <- diag(1, nrow = 3)
d <- c(0,5,0)
A <- matrix(c(-4,-3,0,2,1,0,0,-2,1),3,3) # by column
b <- c(-8,2,0)

# solve the QP
solve.QP(D, d, A, b)

solve.QP(D, d, A, b)$solution

solve.QP(D, d, A, b)$value
```

# 4. Change Problems into Standard Form

original problem

$$
\begin{gather*}
    \max_x f(x) = -0.5x_1^2 - 0.5x_2^2 - 0.5x_3^2 + 4x_2\\
    \text{subject to: }4x_1 + 3x_2 = 8\\
    2x_2 - x_3 \leq 0\\
    2x_1 + x_2 = 2
\end{gather*}
$$

standard form

$$
\begin{gather*}
    -\min_x \{-f(x) = 0.5x_1^2 + 0.5x_2^2 + 0.5x_3^2 - 4x_2\}\\
    \text{subject to: }4x_1 + 3x_2 = 8\\
    2x_1 + x_2 = 2\\
    -2x_2 + x_3 \geq 0
\end{gather*}
$$

problem statement (matrix form)

$$
\begin{equation*}
    D = \begin{bmatrix}
    1 & 0 & 0\\
    0 & 1 & 0\\
    0 & 0 & 1
    \end{bmatrix}; d = \begin{bmatrix}
    0\\
    4\\
    0
    \end{bmatrix}; A = \begin{bmatrix}
    4 & 2 & 0\\
    3 & 1 & -2\\
    0 & 0 & 1
    \end{bmatrix}; b = \begin{bmatrix}
    8\\
    2\\
    0
    \end{bmatrix}; \textbf{meq = 2}
\end{equation*}
$$
```{r}
# problem statement
D <- diag(c(1,1,1))
d <- c(0,4,0)
A <- matrix(c(4,3,0,2,1,0,0,-2,1),3,3)
b <- c(8,2,0)

# solve the problem
solve.QP(D, d, A, b, meq = 2)$solution # optimal solution x

solve.QP(D, d, A, b, meq = 2)$value # minimum value of -f(x)

-solve.QP(D, d, A, b, meq = 2)$value # maximum value of f(x)
```

