---
title: "FE515_2022A_Lecture_6"
author: "Cheng Lu"
output: pdf_document
---
Learning objectives:
  1. Periodic Return
  2. Moments and Sample Moments
    2.1 Mean and Sample Mean
    2.2 Variance and Sample Variance
    2.3 Moments about the Origin and Central Moments
    2.4 Standardized Moments
    2.5 Skewness
    2.6 Kurtosis
    2.7 Other Moments and Sample Moments

# 1. Periodic Return

(Calculation equation and defination in slides)

```{r}
aapl <- read.csv("AAPL.csv", row.names = 1)
head(aapl)
```

calculate log return

```{r}
# adjusted price
aapl.price <- aapl$AAPL.Adjusted # P1,P2,...

# log value
aapl.log.price <- log(aapl.price) # ln(P1),ln(P2),...

# log return
aapl.log.return <- diff(aapl.log.price) # ln(P2)-ln(P1),ln(P3)-ln(P2)...

# convert to simple return
aapl.simple.return <- exp(aapl.log.return) - 1 # (P2/P1)-1,(P3/P2)-1,...
```

distribution of log return

```{r}
# distribution of raw price
hist(aapl$AAPL.Adjusted, 40)

# distribution of log return
hist(aapl.log.return, 40)
```

built-in function for log return

```{r}
library(quantmod)
msft <- getSymbols("MSFT", auto.assign = F)
msft.monthly.return <- periodReturn(msft, period = "monthly")
msft.annual.return <- periodReturn(msft, period = "yearly")
msft.all.return <- allReturns(msft)
msft.all.return
```

subplots of periodic return

```{r}
par(mfrow = c(1,2))
plot(msft.monthly.return)
plot(msft.annual.return)
```
# 2. 2. Moments and Sample Moments

(See slides for defination and calculation equation.)
    
## 2.1 Mean and Sample Mean

```{r}
x <- rnorm(100000, mean = 0, sd = 5)

# sample mean of x
(x.sample.mean <- mean(x)) 
```

## 2.2 Variance and Sample Variance

```{r}
# unadjusted sample variance of x
(x.sample.variance <- mean((x-x.sample.mean)^2)) 

# adjusted sample variance of x
n <- length(x)
(x.sample.variance.adjusted <- n/(n-1)*x.sample.variance) 

# built-in function for adjusted sample variance only
var(x)

# standard deviation
sd(x) ^ 2
```

## 2.3 Moments about the Origin and Central Moments

```{r}
# 3rd sample moment about the origin
(m3.prime <- mean(x^3))

# 4th sample moment about the origin
(m4.prime <- mean(x^4))

# 3rd sample central moment
(m3 <- mean((x-mean(x))^3))
```

## 2.5 Skewness

```{r}
library(moments)
skewness(x)
??skewness
```

## 2.6 Kurtosis
    
```{r}
# kurtosis
kurtosis(x) - 3

?kurtosis
# excess kurtosis
PerformanceAnalytics::kurtosis(x)
```
    
## 2.7 Other Moments and Sample Moments

```{r}
cov(rnorm(100), rnorm(100))
```

