---
title: "FE515_2022A_Lecture_3"
author: "Cheng Lu"
output: pdf_document
---

Learning objectives:
  1.Date and Time
    1.1 Date Object
    1.2 Date and Time Objects
  2. Visualization
    2.1 X-Y Plot
    2.2 Scatter Plot
    2.3 Bar Plot
    2.4 Box Plot

# 1. Date and Time

## 1.1 Date Object

as.Date() function convert character to objects of class ”Date” representing calendar dates.

```{r}
myDate <- as.Date("2022-06-07")
class(myDate)
mode(myDate)

# internally it is number of days since Jan 01, 1970
as.numeric(myDate)
myDates <- as.Date(c("1970-01-01", "1970-01-02"))
as.numeric(myDates)
```

as.Date() can only recognize some standard time type.

```{r}
# standard input
as.Date("1970/01/01")

# non-standard input
as.Date("01/01/1970") 

# set up date format when we generate dates
as.Date("02/24/2015", format = "%m/%d/%Y")
as.Date("15,Feb-24", format = "%y,%b-%d")
as.Date("February24 2015", format = "%B%d %Y")
```

format() function convert `Date` object to character.

```{r}
# convert system date to character
date.1 <- Sys.Date()
class(date.1)

date.1.convert <- format(Sys.Date()) 
date.1.convert
class(date.1.convert)

# specify conversion format
myDate2 <- format(Sys.Date(), "%Y/%m/%d")
myDate2

?as.Date
```

Other useful functions for ”Date” objects*

```{r}
myDate <- as.Date("1970-01-01")

# get date property
months(myDate)
quarters(myDate)
weekdays(myDate)

# date sequence
myDate + 1
seq(from = myDate, to = myDate + 31, by = "day")
seq(myDate, myDate + 31, by = "week")
```

## 1.2 Date and Time Objects

The POSIXct and POSIXlt classes allow for `dates` and `times` with control for `time zones`

- The POSIXct (POSIX standards of calendar time) class stores date/time values as the `number of seconds` since January 1, 1970
- The POSIXlt (POSIX standards of local time)class stores them as a list with elements for second, minute, hour, day, month, and year, among others. 
- Unless you need the list nature of the POSIXlt class, the POSIXct class is the usual choice for storing dates in R

```{r}
myDateTime1 <- as.POSIXct("01,01,2014 10:20:20", format = "%d,%m,%Y %H:%M:%S")
myDateTime2 <- as.POSIXlt("01,01,2014 10:20:20", format = "%d,%m,%Y %H:%M:%S")
myDateTime3 <- as.POSIXlt("01,02,2014 10:21:20", format = "%d,%m,%Y %H:%M:%S")

class(myDateTime1)
class(myDateTime2)

mode(myDateTime1)
mode(myDateTime2)
```

strptime() and strftime()

```{r}
# convert string to date object 
# it's the core part of as.POSIXlt()
dt1 <- strptime("2013.12/23 01:00:34", "%Y.%m/%d %H:%M:%S")
dt2 <- strptime("2013.12/23 01:02:37", "%Y.%m/%d %H:%M:%S")

dt1
class(dt1)

# convert date object to string
dt1 <- strftime(dt1, "%Y.%m/%d %H:%M:%S")
```

operation on date and time object

```{r}
# get date property
dt1$sec # other elements: min, hour, mday, wday, yday, mon, year

# difference in date object 
dt2 - dt1

myDateTime1 - myDateTime3
myDateTime2 - myDateTime3

difftime(dt2, dt1, units = "secs")
```

# 2. Visualization

## 2.1 X-Y Plot

plot() function is the most basic graphic function in R.
It doesn’t need any package to start.
Lines() function add line segments to the plot.
Points() function add points to the plot.

```{r}
x <- seq(0, 5, by = 0.5)
y <- sin(x)
y2 <- cos(x)

{
  plot(x, y, 
     xlim = c(-1,6), 
     ylim = c(-1.5, 1.5),
     main = "Title", 
     xlab = "x label", 
     ylab = "y label", 
     col = rgb(0,1,0,0.5))

# add one addtional line
lines(x, y2, type = "c", col = "red")
}

```

main: title of the plot
xlab,ylab: lable for x or y axis
xlim,ylim: value range on x or y axis
lwd: line width
lty: type of line
pch: type of point
col: color for line or point

## 2.2 Scatter plot

Scatter Plot of Miles per gallon against weight (weight in x-axis and Miles per gallon in y-axis

```{r}
plot(mtcars$wt, mtcars$mpg, 
     main = "Scatter Plot",
     xlab = "Weight", 
     ylab = "Miles per gallon")

```


## 2.3 Bar plot

### 2.3.1 discrete data

Barplot of frequency against number of cylinders

```{r}
?mtcars
table(mtcars$cyl)
barplot(table(mtcars$cyl),xlab = "Number of cylinders", ylab = "Frequency")
```

### 2.3.2 continuous data*

Use cut() function divide data into intervals, 
then use table() function to generate tables,
then use barplot() to plot.

```{r}
cut(mtcars$wt, breaks = 1:6) # also use "breaks = 6" to get 6 equal lenght intervals

table(cut(mtcars$wt, breaks = 1:6))

barplot(table(cut(mtcars$wt, breaks = 1:6)),xlab = "weight", ylab = "Frequency")
```

## 2.4 Box Plot

fivenum() finds the minimum, first quartile, median, third quartile, maximum. And boxplot()
summarize these numbers in a plot.

```{r}
fivenum(mtcars$hp)
# setup 2 figures in 1 row and 2 column which filled by rows

par(mfrow = c(1, 2)) # how about mfcol?
boxplot(mtcars$mpg, xlab = "Mile per gallon")
boxplot(mtcars$hp, xlab = "Gross horsepower")
```

