---
title: "FE515_2022A_Lecture_4"
author: "Cheng Lu"
output: pdf_document
---
Learning objectives:
  1. quantmod package
  2. Historical Prices
  3. Quoted Prices
  4. Option prices
  5. Output file

# 1. quantmod package

download and import package
```{r}
#install.packages("quantmod")
library(quantmod)
```

download historical stock price
```{r}
getSymbols(Symbols = "MSFT")
MSFT <- data.frame(MSFT)
```

overview of data
```{r}
head(MSFT, 3)
```

another example
```{r}
ticker.list <- c("SPY","AAPL")
data.env <- new.env()
getSymbols(Symbols = ticker.list, 
           from = "2001-01-01", 
           to = "2001-01-31",
           auto.assign = T,
           env = data.env
           )

head(data.env$AAPL) # show first few observations
tail(data.env$AAPL) # show last few observations

# The default source is "yahoo", which is the only source now
```

# 2. Visualization

plot stock price series
```{r}
chartSeries(MSFT)
```

subsetting
```{r}
chartSeries(MSFT, subset='2007-01::2007-03')
```

change color(Candlestick Charts)
```{r}
chartSeries(AAPL,theme=chartTheme('white'))
```

technical indicators
```{r}
#no volume
chartSeries(MSFT,TA=NULL) 

#add volume and Bollinger Bands from TTR
chartSeries(MSFT,TA=c(addVo(),addBBands()), subset='2007-01::2007-03')
addMACD()
```

# 3. Quoted Prices

current quote price
```{r}
tail(MSFT)
getQuote("MSFT")
```

# 4. Option prices*

download option price

Option is a contract that a investor has right to buy or sell the underlying asset at specific time (expiration date or maturity) with specific price (strike price)
```{r}
AAPL.option <- getOptionChain("AAPL")

# a named list
mode(AAPL.option)

# two elements
names(AAPL.option)# use names() to check variables of lists

# data type of each element
is.data.frame(AAPL.option)
is.data.frame(AAPL.option$calls)
```

understand option data

- The first observation is a call option that a investor has right buy 100 share of “AAPL” on “year-month-day” with price $##.## (Strike)
- The most recent traded (last quoted) price for the contract is $##.## (Last)
- The highest price a buyer willing to pay (bid price) for the option is $##.## (Bid)
- The lowest price a seller willing to sell (ask price) for the option is $##.## (Ask)
- The price will change during trading day when investors trade the option

```{r}
head(AAPL.option$calls)
```

change expiration date
```{r}
# set expiration date
AAPL.option1 <- getOptionChain("AAPL", Exp = "2022-10-14")
names(AAPL.option1)

# multiple expiration date
AAPL.options <- getOptionChain("AAPL", c("2022-10-07","2022-10-14"))
names(AAPL.options) # expired between "2022-07-01" and "2022-07-15"
names(AAPL.options$Oct.07.2022)
```

more expiration dates
```{r}
# all options expired in 2022 and 2023
AAPL.options2 <- getOptionChain("AAPL", "2022/2023")
names(AAPL.options2) 

# all options available
AAPL.options.all <- getOptionChain("AAPL",NULL)
names(AAPL.options.all)

# strike price of call options expired on "2022-10-07"
AAPL.options.all$Oct.07.2022$calls$Strike
```

calculate option price
```{r}
names(AAPL.options.all$Oct.07.2022$calls)
AAPL.options.all$Oct.07.2022$calls$Price <- 0.5*(AAPL.options.all$Oct.07.2022$calls$Bid
                                              + AAPL.options.all$Oct.07.2022$calls$Ask)
names(AAPL.options.all$Oct.07.2022$calls)
```

**more details about option price calculations on slides**

# 5. Output file*

output as a csv file
```{r}
file.path <- "data2022_10_03Exp2022_10_07calls.csv"
write.csv(AAPL.options.all$Oct.07.2022$calls, file.path)
```


output as a Rdata file
```{r}
file.path <- "data2022_10_03Exp2022_10_07calls.Rdata"
x <- AAPL.options.all$Oct.07.2022$calls
save(x, file = file.path)
```

output as a txt file
```{r}
file.path <- "data2022_10_03Exp2022_10_07calls.txt"
write.table(AAPL.options.all$Oct.07.2022$calls, file = file.path)
```

