---
title: "FE515_2022A_Lecture_5"
author: "Cheng Lu"
output: pdf_document
---
Learning Objectives:
  1. Coin Filps
  2. Normal Distribution
  3. Generate Random Number

# 1. Coin Filps

sample() function take samples of the specified size from the elements of x using either with
or without replacement.

```{r}
# take samples from population with replacement
sample(x=c(1,2,3),size = 1, replace = T)

# take samples from population without replacement
sample(x=c(1,2,3),size = 2,replace = F)
```

Coin flips example

```{r}
no.heads <- 0
result.vec <- NULL # vector contains all probabilities

tmp <- sample(x = c(1, 0), size = 1, replace = T, prob = c(0.5, 0.5))

# add tmp to number of head
no.heads <- no.heads + tmp
result.vec <- c(result.vec, no.heads/1)
```


```{r}
no.heads <- 0
result.vec <- NULL # vector contains all probabilities

for (flips in 1:1000){
# x = c(1, 0), 1 means head, 0 means tail
tmp <- sample(x = c(1, 0), size = 1, replace = T, prob = c(0.5, 0.5))

# add tmp to number of head
no.heads <- no.heads + tmp
result.vec <- c(result.vec, no.heads/flips)
}

no.heads
result.vec
plot(result.vec)
```

Second implementation

```{r}
tmp <- sample(x = c(0,1), size = 1000, replace = T, prob = c(0.5,0.5))
no.heads <- cumsum(tmp)
result.vec <- no.heads/1:1000
plot(result.vec)
```

Visualization

```{r}
plot(1:1000, result.vec, type="l", main="Coin Flips", xlab='flip time', ylab='probability')
```

A founction of `face up` probability

```{r}
coinFlip <- function(head.prob) {
no.heads <- 0
result.vec <- NULL
for (flips in 1:1000){
  tmp <- sample(x=c(1, 0), size=1, replace=T,prob=c(head.prob, 1 - head.prob))
  no.heads <- no.heads + tmp
  result.vec <- c(result.vec, no.heads/flips)
}
plot(1:1000, result.vec, type="l", ylim = c(0, 1))
}
```

Change `face up` probability

```{r}
coinFlip(0.5)
coinFlip(0.7)
coinFlip(0.9)
coinFlip(1)
```

# 2. Normal Distribution

sample from  normal distribution

$$
X \sim N(\mu, \sigma)
$$

```{r}
x <- rnorm(n = 10000, mean = 0, sd = 1)
hist(x) # histogram
hist(x, nclass = 40, main = "mu = 0, sigma = 1")
```

Change variance

```{r}
x <- rnorm(n = 10000, mean = 0, sd = 5)
hist(x, nclass = 40, main = "mu = 0, sigma = 5")
```

Control randomness

```{r}
set.seed(888888)
rnorm(5)
```

## 2.1 Probability density function (PDF) of normal distribution

$$
f(x)=\frac{1}{\sqrt{2 \pi} \sigma} e^{-\frac{(x-\mu)^{2}}{2 \sigma^{2}}}
$$

```{r}
dnorm(x = 0)
```

## 2.2 Cumulative distribution function (CDF) of normal distribution

$$
F(x)=P(X \leq x)=\int_{-\infty}^{x} f(t) d t
$$

```{r}
pnorm(q = 0) # F(0) = P(X <= 0) = 0.5
```

## 2.3 Quantile function of normal distribution

$$
x=F^{-1}(p)
$$

```{r}
qnorm(p = 0.5)
```

## 2.4 Other distribution*

Poisson distribution

$$
f(k ; \lambda)=\operatorname{Pr}(X=k)=\frac{\lambda^{k} e^{-\lambda}}{k !}
$$
```{r}
x <- rpois(1000, lambda = 2)
hist(x, nclass = 40)
```

Exponential distribution

$$
f(x ; \lambda)= \begin{cases}\lambda e^{-\lambda x} & x \geq 0 \\ 0 & x<0\end{cases}
$$

```{r}
x <- rexp(1000)
hist(x, nclass = 40)
```

Student's t-distribution

```{r}
set.seed(1)
x <- rt(1000, df = 10)
hist(x, nclass = 40)
```

# 3. Generate Random Number

$$
X_{n+1}=\left(a * X_{n}+b\right) \bmod m
$$
 - $m$ is modulus
 - $a$ is multiplier
 - $b$ is increment
 - $0 \leq X_{0}<m$ is seed
 - Park and Miller suggests $m=2^{31}-1, a=7^{5}, b=0$.

```{r}
# hyper-parameter
seed <- 1 # let the seed be 1
n <- 5 # quantity of random numbers
m <- 2 ^ 31 - 1
a <- 7 ^ 5
b <- 0
x <- rep(NA, n)

# initial value
x[1] <- (a * seed + b) %% m

# random number sequence
for(i in 1:(n-1)){
x[i + 1] <- (a * x[i] + b) %% m
}
x

seed <- x[n] # change the seed
x/m
```

Wrap up everthing into a function

```{r}
# generate random number from uniform distribution
seed <- 1 # let the seed be 1 in global environment

rnd <- function(n){
m <- 2 ^ 31 - 1
a <- 7 ^ 5
b <- 0
x <- rep(NA, n)
x[1] <- (a * seed + b) %% m
for(i in 1:(n-1)){
x[i + 1] <- (a * x[i] + b) %% m
}
seed <<- x[n] # change the seed in global environment
return(x/m)
}
```

Test the function

```{r}
rnd(5) # the first few numbers are usually bad, we can discard them
rnd(5)

# use the same seed to replicate the results
seed <- 1
rnd(5) # same number for seed = 1

# change the seed
seed <- 100 
rnd(5)

# use the current system time as seed
seed <- as.numeric(Sys.time())
rnd(5)
```

Visualize the resulting distribution

```{r}
U <- rnd(10000) # 10000 uniform r.v.
hist(U)
```

```{r}
# qnorm() is the quantile function (inverse CDF) of normal distribution
X <- qnorm(U) # X follows normal distribution
hist(X, nclass = 40) # histogram of normal distribution
hist(rnorm(10000), 40) # equivalent to the above statement
hist(qexp(rnd(10000), rate = 1), nclass = 40) # exponential distribution
```

## 3.1 Example*

Calculate the value of pi (see slides for details)

```{r}
N <- 10000
x <- rnd(N)
y <- rnd(N)

n.red.bool <- x^2 + y^2 <= 1
n.red.num <- sum(n.red.bool) # vectorized calculation

area.quarter.circle <- n.red.num / N
(Pi <- 4 * area.quarter.circle) # force output
```

