---
title: "FE515_Lecture_12"
author: "Cheng Lu"
output: pdf_document
---
Learning objectives:
  1. Interpolation
    1.1 Linear Interpolation
    1.2 Spline Interpolation

  2. Numerical Integration
    2.1 Trapezoidal rule
    2.2 Simpson's rule

# 1. Interpolation

## 1.1 Linear Interpolation

```{r}
set.seed(1)
n <- 9
x <- 1:n # x = 1, 2, 3, ..., 9
y <- rnorm(n) # generate y from standard normal
plot(x, y)
lines(x, y, col = "red")
```

find y at point x = 4.4

$$
y_{o u t}=y_{4}+\frac{y_{5}-y_{4}}{x_{5}-x_{4}}\left(x_{o u t}-x_{4}\right)
$$

```{r}
# interpolation at x = 4.4
xout <- 4.4
which(x < xout) # find indices for x < x_out
max(which(x < xout)) # max index
id <- max(which(x < xout)) # id <- max index
yout <- y[id] + (y[id+1] - y[id])/(x[id+1] - x[id])*(xout - x[id])

# visualize
plot(x, y)
lines(x, y, col = "blue")
points(xout, yout, col = "red")
```

interpolate at many points

```{r}
xout <- c(1.1, 1.3, 2.5, 4.4, 5.6, 8.8)
yout <- NULL

# Loop
for(i in 1:length(xout)){
id <- max(which(x < xout[i]))
yout[i] <- y[id] + (y[id+1] - y[id])/(x[id+1] - x[id])*(xout[i] - x[id])
}

# visualize
plot(x, y)
lines(x, y, col = "blue")
points(xout, yout, col = "red")
```

vectorized version

```{r}
# Vectorized
func1 <- function(xout){
id <- max(which(x < xout))
yout <- y[id] + (y[id+1] - y[id])/(x[id+1] - x[id])*(xout - x[id])
}
yout <- sapply(xout, func1)

# visualize
plot(x, y)
lines(x, y, col = "blue")
points(xout, yout, col = "red")
```

use built-in function

```{r}
# method 1
yout <- approx(x,y,xout = xout)$y

# visualize
plot(x, y)
lines(x, y, col = "blue")
points(xout, yout, col = "red")
```

second method

```{r}
# method 2
yout.fun <- approxfun(x, y, method = 'linear')# function
yout <- yout.fun(xout)

# visualize
plot(x, y)
lines(x, y, col = "blue")
points(xout, yout, col = "red")
```

## 1.2 Spline Interpolation

$$
A=\left[\begin{array}{cccccc}
1 & x_{1}-x_{1} & \left(x_{1}-x_{1}\right)^{2} & \left(x_{1}-x_{1}\right)_{+}^{3} & \ldots & \left(x_{1}-x_{n-1}\right)_{+}^{3} \\
1 & x_{2}-x_{1} & \left(x_{2}-x_{1}\right)^{2} & \left(x_{2}-x_{1}\right)_{+}^{3} & \ldots & \left(x_{2}-x_{n-1}\right)_{+}^{3} \\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots \\
1 & x_{n}-x_{1} & \left(x_{n}-x_{1}\right)^{2} & \left(x_{n}-x_{1}\right)_{+}^{3} & \ldots & \left(x_{n}-x_{n-1}\right)_{+}^{3} \\
0 & 0 & 2 & 0 & \ldots & 0 \\
0 & 0 & 2 & 6\left(x_{n}-x_{1}\right) & \ldots & 6\left(x_{n}-x_{n-1}\right)
\end{array}\right] ; \theta=\left[\begin{array}{c}
a \\
b \\
c \\
d_{1} \\
\vdots \\
d_{n-1}
\end{array}\right] ; z=\left[\begin{array}{c}
y_{1} \\
y_{2} \\
\vdots \\
y_{n} \\
0 \\
0
\end{array}\right]
$$

$$
A \theta=z
$$
```{r}
theta
```


```{r}
A <- cbind(1, x-x[1], (x-x[1])^2, pmax(outer(x, x[-n], "-"), 0)^3) # outer(x,x,"-") generates matrix M with M_ij = (x_i - x_j)
A <- rbind(A, 0, 0) # initialize last 2 row
A[c(n+1,n+2), 3] <- 2
A[n+2, 4:(n+2)] <- 6*(x[n] - x[-n]) # last row
z <- c(y, 0, 0)
theta <- solve(A,z) # solve linear system

# set up the model
a <- theta[1]
b <- theta[2]
c <- theta[3]
d <- theta[-c(1,2,3)]
yout <- as.vector(a + b*(xout-x[1]) + c*(xout-x[1])^2 + pmax(outer(xout, x[-n], "-"), 0)^3%*%d)

# visualize
plot(x, y)
lines(x, y, col = "blue")
points(xout, yout, col = "red")
```

use built-in function

```{r}
set.seed(1)
# method 1
yout.s <- spline(x,y, xout = xout, method = "natural")
yout <- yout.s$y

# method 2
yout.s.fun <- splinefun(x, y, method = "natural")
yout.s.fun(xout)

# visualize
plot(x, y)
lines(seq(1,9,by=0.1),yout.s.fun(seq(1,9,by=0.1)), col = "blue")
points(xout, yout.s.fun(xout), col = "blue")
```

# 2. Numerical Integration

# 2.1 Trapezoidal rule

How a trapezoid looks like

```{r}
# Define the coordinates of the trapezoid's vertices
trapezoid_x <- c(0.3, 0.7, 0.9, 0.1, 0.3)
trapezoid_y <- c(0.3, 0.3, 0.7, 0.7, 0.3)

# Set up the plotting window
plot(NULL, type = "n", xlim = c(0, 1), ylim = c(0, 1), axes = FALSE, xlab = "", ylab = "")

# Draw the trapezoid using the polygon() function
polygon(trapezoid_x, trapezoid_y, col = "blue", border = "black")
```


Consider the following example

$$
f(x) = (x^3 - 8) \cdot \phi(x)
$$


```{r}
f <- function(x)(x^3-8)*dnorm(x) # define integrand function
a <- 2 # lower bound of x
b <- 10 # upper bound of x
curve(f, from = a, to = b)
```

apply trapezoidal rule to approximate the area under the curve

```{r}
# Define the function
f <- function(x) (x^3 - 8) * dnorm(x)

# Define the integration interval and trapezoid vertices
a <- 2.5
b <- 3.5
trapezoid_x <- c(a, b, b, a)
trapezoid_y <- c(f(a), f(b), 0, 0)

# Set up the plotting window
plot(f, type = "n", xlim = c(2, 10), ylim = c(0, 0.2), xlab = "x", ylab = "f(x)")

# Draw the curve of the function
curve(f, add = TRUE, col = "blue", lwd = 2)

# Draw the trapezoid using the polygon() function
polygon(trapezoid_x, trapezoid_y, col = rgb(0.8, 0.8, 0.8, 0.5), border = "black")

# Add labels for the trapezoid vertices
text(trapezoid_x[1], trapezoid_y[1], paste("(", a, ",", round(f(a), 2), ")", sep = ""), pos = 3)
text(trapezoid_x[2], trapezoid_y[2], paste("(", b, ",", round(f(b), 2), ")", sep = ""), pos = 3)

```

repeat over all intervals

```{r}
n <- 1000
x <- seq(a, b, length.out = n+1)
(b - a)/n*(sum(f(x)) - 0.5*f(a) - 0.5*f(b))
```

# 2.2 Simpson's rule


```{r}
# Define the function
f <- function(x) (x^3 - 8) * dnorm(x)

# Define the integration interval and the number of subintervals
a <- 2.5
b <- 3.5
n <- 2

# Calculate the width of the subintervals
h <- (b - a) / n

# Calculate the x values for each subinterval
x_values <- seq(a, b, by = h)

# Set up the plotting window
plot(f, type = "n", xlim = c(2, 4), ylim = c(-0.2, 0.2), xlab = "x", ylab = "f(x)")

# Draw the curve of the function
curve(f, add = TRUE, col = "blue", lwd = 2)

# Draw the parabolic segments and shaded areas
for (i in seq(1, n, by = 2)) {
  x1 <- x_values[i]
  x2 <- x_values[i + 1]
  x3 <- x_values[i + 2]

  y1 <- f(x1)
  y2 <- f(x2)
  y3 <- f(x3)

  parabola_points <- get_parabola_points(x1, y1, x2, y2, x3, y3)
  polygon(c(parabola_points$x, rev(parabola_points$x)), c(parabola_points$y, rep(0, length(parabola_points$y))), col = rgb(0.8, 0.8, 0.8, 0.5), border = NA)

  lines(parabola_points$x, parabola_points$y, col = "red", lwd = 2)
}

# Function to get the points for drawing the parabolic segments
get_parabola_points <- function(x1, y1, x2, y2, x3, y3) {
  t <- seq(0, 1, length.out = 100)
  x <- x1 * (1 - t)^2 + 2 * x2 * t * (1 - t) + x3 * t^2
  y <- y1 * (1 - t)^2 + 2 * y2 * t * (1 - t) + y3 * t^2
  return(data.frame(x = x, y = y))
}

# Draw vertical lines and points at the nodes
for (i in 1:(n + 1)) {
  x <- x_values[i]
  y <- f(x)
  points(x, y, col = "black", pch = 19)
  segments(x, y, x, 0, col = "black", lty = 2)
}

# Add labels for the nodes
for (i in 1:(n + 1)) {
  x <- x_values[i]
  y <- f(x)
  text(x, y, paste("(", round(x, 1), ",", round(y, 2), ")", sep = ""), pos = 3)
}

```

```{r}
f <- function(x)(x^3-8)*dnorm(x) # define integrand function
a <- 2 # lower bound for x
b <- 10 # upper bound for x
n <- 1000
x <- seq(a, b, length.out = n+1)
as.vector((b - a)/(3*n)*c(1,rep(c(4,2),n/2)[-n],1)%*%f(x))
```

# 2.3 Using built-in functions

```{r}
f <- function(x)(x^3-8)*dnorm(x)# define integrand function
integrate(f, 2, 10)
s <- integrate(f, 2, 10)
s$value
```

